Notes and other text

### 2023-04-14 doc/manuscript/translocation.qmd
Address comments on population viability analyses

Comparison of genomes of frogs from Bd-naive versus recovering populations (described above) indicated substantial differences in allele frequencies and genomic regions, consistent with strong selection following Bd exposure.
In this section, we describe the outcome of frog reintroductions in which animals for reintroduction were collected from recovering populations, including two of the five recovering populations in the frog evolution study (@fig-allelefrequencies A: population 1 and 4).
Reintroductions of *R. sierrae* were conducted as part of an effort to reestablish populations of this endangered species.
These reintroductions were conducted in Yosemite National Park (California, USA) and involved the collection of adult frogs from large, Bd-positive donor populations that have recovered in recent years following Bd-caused declines [@knapp2016] and translocating them to nearby recipient sites [@joseph2018].
We refer to these reintroductions as "translocations" [@seddon2014].
During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of frogs to `r n_distinct(d1$site_id)` recipient sites (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap SI).
Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs (median = `r transfrogs_median`) collected from 1 of `r n_distinct(d1$donor)` donor populations (@fig-yosemap SI).
Estimates of frog survival during the one year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-survival-earlylate SI), which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads changed little between immediately prior to translocation versus during the one year after translocation (@fig-bdload-beforeafter SI), and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale) [@vredenburg2010; @joseph2018].

@@

In addition to survival......

<!--# To allow comparison with other translocation studies, adding information about the proportion of translocated populations at which evidence of successful reproduction was observed in the 1-2 years following translocation. -->

@@

We then included demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of juvenile recruitment to the adult class $\omega$ (the transition that we expect to be the most subject to environmental stochasticity in the presence of Bd) into our model described above and used it to predict long-term population viability.
<!--# In regard to the previous sentence, is it worth stating that the inclusion of demographic stocahsticity and environmental variability will not change estimates of lambda? That might not be apparent. --><!--# In the following sentences, I edited the text to present the results based on omega = 0.3 first and then the results with reduced omega of 0.15. This matches the order in which we present the results in the previous paragraph. -->Our model predicted that, following a single introduction of 40 adult individuals into a population and using the baseline value for $\omega$ of 0.3, all six populations with the highest adult survival probabilities had 50-year extinction probabilities less than 0.5 (<!--# provide actual range here -->@fig-viability B).
This indicates strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment (@fig-viability B).
When we decreased $\omega$ from 0.3 to our most conservative estimate of 0.15, 50 year extinction probability remained less than 0.5 for three populations and was 0.5 to 0.X <!--# give actual range here -->for three additional populations (@fig-viability B).
In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history (@fig-viability-supp B; population 70550, surveyed for 16 years).
In summary, our model demonstrates that given observed yearly adult survival probabilities of translocated frogs, 50% of our translocated populations have a high probability of population growth and long-term viability in the presence of Bd.
This is likely a conservative estimate because there is evidence that naturally-recruited adults have higher survival probability than translocated adults (@fig-compare_surv_probs SI), but we considered these probabilities to be equal in all but three of our populations where we had sufficient data to distinguish these different probabilities.
Finally, five of the seven populations translocated after 2013 were predicted to have long-term viability, compared to only one of five populations translocated prior to 2013.
Thus, there is some trend in our improved ability to select appropriate habitat to promote long-term viability of translocated, resistant populations.

@@

<!--# If we switch the order of sections in the Results, we might want to mirror that order here too: translocations > selection results & evolutionary rescue > viability modeling -->

@@

We provide three lines of evidence that support evolutionary rescue as a primary mechanism of recovery in MYL frog systems.
First, our genomic analyses comparing naive and recovering MYL frog populations found notable differences in allele frequencies between adjacent Bd-naive and recovering <!--# Check on using "persistent" vs "recovering" vs "recovered" --> <!--# Roland response: Use "recovering" throughout -->populations, consistent with strong selection.
Importantly, some of the regions of the genome that showed signatures of selection are associated with known cellular and immunological mechanisms contributing to disease resistance in other species generally [@robledo2018; @robledo2020] and amphibians specifically [@rosenblum2008; @ellison2014] (for more information, see SI X).
For example, we identified a \~940kb region on contig 8 that shows evidence of selection in recovering populations and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate numerous previous studies linking MHC genes to Bd survival [e.g., @savage2011; @bataille2015].
Broadly, our genomic analyses imply that selection and rapid evolutionary change has contributed to survival in MYL frogs in the presence of Bd.
<!--# For brevity in the discussion, I tried to a pull a few key bullets out of Allie's text. Allie and Bree: Edit and expand as you see fit. -->

The second line of evidence that evolutionary rescue is a strong candidate mechanism for recovery is that repeated translocations and long-term monitoring of naturally recovered populations have demonstrated that resistant phenotypes are maintained beyond the natal environment in the presence of Bd (as reflected by the high year-one adult survival in 6 out of 12 translocated populations).
Moreover, resistance has a quantifiable phenotype in this system -- resistant, persisting populations have Bd infection intensities below the level known to cause significant host mortality [@vredenburg2010, @knapp2016].
Translocated populations maintain infection intensities below this threshold of infection intensity such that infection intensity is not a significant predictor of adult survival following translocations to new environments.
Third, high adult survival in the presence of Bd persists into subsequent generations in non-natal environments, suggesting that resistance is heritable to some degree.
Although previous studies have documented signatures of selection in hosts following disease-induced declines [@epstein2016, @gignoux-wolfsohn2021], our study is unique in that these three lines of evidence provide a strong case that evolved resistance begets population recovery in both natal and non-natal habitats, consistent with evolutionary rescue.

@@

Additionally, we identified a region on contig 19 that had much lower in the recovering group and consisted of one gene -- interferon-induced very large GTPase 1-like (GVINP1).
Interferons boost immune system processes and previous studies have linked interferon related genes to persisting post-Bd [@byrne2020] or increased expression when exposed to Bd in the lab [@rosenblum2008; @ellison2014].
Variants in the specific gene identified here, GVINP1, explained \~20% of the resistance to Amoebic Gill Disease in Atlantic Salmon, a finding confirmed via expression analysis [@robledo2018; @robledo2020].
Furthermore, a key difference in susceptibility between populations of alpine tree frogs (*Litoria verreauxii alpina*) was the lower expression of IFN-induced very large GTPase in the more resistant, long-exposed population and upregulation of this same gene in more susceptible populations [@grogan2018].
Our study therefore contributes to the growing body of evidence linking this gene to disease resistance in general, and specifically to Bd survival in amphibians.
Finally, we found evidence of high FST and higher in recovering populations in the complement genes C6 and C7 <!--# It is clear why we talk about the other two regions (the ones highlighted in figure 2) and RIN3 - see comments above. But I guess the only reason this is interesting is because it is a complement region and it shows balancing rather than directional selection. If it feels like we can't justify discussing it we can cut it and just focus on the other two.  -->.
This could indicate that balancing selection is acting in this region of the genome, favoring a diverse set of alleles in the recovering group which has been shown for C6 in humans [@soejima2005].
Overall, key outlier genes in our dataset link directly to experimental and observational studies of immune responses to pathogens -- and often specifically to Bd responses in amphibians -- adding confidence to our conclusion that selection and rapid evolutionary change has contributed to survival in MYL frogs.

@@

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ ("$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year $p_F$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile $p_{J1}$.
The parameter $\omega$ gives the probability of a juvenile successfully recruiting to an adult.
<!--# Can the omega parameter be added to the list of other parameters: "..., probability of a first-year juvenile remaining as a juvenile pj1, and probability of a juvenile successfully recruiting to an adult omega." Listing omega separately seems unnecessary, at least to me. -->

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated frogs. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (given by the dashed line). (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505. <!--# Remove period after plot labels, e.g., "A", not "A.". -->](figures/pop_viability_figures_for_manuscript_v2.jpg){#fig-viability}

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$, $\omega$, and *F (*number of eggs produced by a female frog in a year that successfully hatch) <!--# Necessary to define F (when sigma and omega are not defined)? Maybe instead just cite "Table 2 SI"?) -->. Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

<!--# Remove period after plot labels -->

<!--# Left panel color palette is 12-color color-blind friendly, but is not the most friendly color palette to look at.  We can discuss. Roland says: Maybe try a palette that you like better and mention that order of bars (left to right) is the same as in the legend (top to bottom). In addition, if you think it is necessary, for those colors that will be hard for eyeballs like mine to distinguish, we can either tweak the colors or add diagnonal lines (ggpattern::geom_bar_pattern). -->

@@

![Comparison of the average yearly adult survival probabilities for adults translocated to each of three sites versus adults that were naturally recruited at each site (as a result of reproduction by translocated frogs). In contrast to @fig-translocation-survival, these are not survival probabilities from the first year following translocation, but instead represent averaged survival probabilities across multiple years. Points are median estimates and error bars give the 90% uncertainty intervals around the estimates. All estimates are derived from the mrmr package. <!--# Use 95% uncertainty intervals for consistency? Also, use dark blue instead of gray for naturally-recruited? Might show up better. -->](figures/compare_surv_probs.jpg){#fig-compare_surv_probs}

@@

<!--# How to insert >= symbol in header and table cells> -->

```{r}
#| label: tbl-survival-earlylate
#| tbl-cap: "Proportion of populations translocated early versus late in the study period (< 2013 or >= 2013, respectively) for which the probability of survival was < 0.5 or >= 0.5."
#| echo: FALSE
#| message: FALSE

viability_table <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
  distinct(site_id, year) %>% 
  group_by(site_id) %>% 
  filter(year == min(year)) %>% 
  mutate(site_id = as.character(site_id),
         survival = if_else(site_id %in% c(70414, 70370, 70134, 70641, 70619, 70505), "< 0.5", ">= 0.5"),
         period = if_else(year < 2013, "early", "late")) %>% 
  group_by(period) %>% 
  count(period, survival) %>% 
  mutate(n_period = sum(n), 
         prop = (n / n_period)) %>% 
  mutate(across(prop, round, 3)) %>% 
  arrange(period, survival)

  knitr::kable(viability_table, format = "pipe", align = "lcccc", col.names = c("Period", "Survival probability", "Number of translocations", "Total translocations", "Proportion of translocations")) 
```

>>>>

<!--# Bree comment: Here or elsewhere - it feels like we need a more specific bridge from the genomics section into the translocation section. I think we need to be explicit about how the populations we chose for the genomics section relate to the populations used in translocations. Even if they aren't matched in any way, we should reiterate something about how the experimental designs comport. For example, we could explain how the genomic analyses were conducted with multiple recovering populations so we know evidence for selection is strong in multiple areas of the genome in multiple recovering populations. Roland response: I added what I hope are bridging sentences between all three sections. -->

Comparison of genomes of frogs from Bd-naive versus recovering populations (described above) indicated substantial differences in allele frequencies and genomic regions, consistent with strong selection following Bd exposure.
In this section, we describe the outcome of frog reintroductions in which animals for reintroduction were collected from recovering populations, including two of the five recovering populations in the frog evolution study (@fig-allelefrequencies A: population 1 and 4).
<!--# Check this --> Reintroductions of *R. sierrae* were conducted as part of an effort to reestablish populations of this endangered species.
These reintroductions were conducted in Yosemite National Park (California, USA) and involved the collection of adult frogs from large, Bd-positive donor populations that have recovered in recent years following Bd-caused declines [@knapp2016] and translocating them to nearby recipient sites [@joseph2018].
We refer to these reintroductions as "translocations" [@seddon2014].
During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of frogs to `r n_distinct(d1$site_id)` recipient sites (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap SI).
Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs (median = `r transfrogs_median`) collected from 1 of `r n_distinct(d1$donor)` donor populations (@fig-yosemap SI).
Estimates of frog survival during the one year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival).
As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in subsequent translocations to that site (@fig-transsurvival-postdens SI).
This suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival.
In addition, 1-year survival was higher for frogs translocated later in the study period than earlier (@tbl-viability-earlylate SI) <!--# Mention of viability is confusing because this issue isn't dealt with until the third section. Perhaps make the categorization based on frog survival? -->, which we suggest resulted primarily from our improved ability over time to choose recipient sites with higher habitat quality for *R. sierrae*.
Finally, Bd loads changed little between immediately prior to translocation versus during the one year after translocation (@fig-bdload-beforeafter SI), and loads were nearly always well below the level indicative of severe chytridiomycosis and associated frog mortality (5.8 ITS copies on a log~10~ scale) [@vredenburg2010; @joseph2018].
<!--# Add median Bd load, proportion below 5.8 copies threshold? --> <!--# The Methods-Translocation section contains text that repeats similar results. Edit to reduce repetition. --> <!--# To allow comparison with other translocation studies, adding information about the proportion of translocated populations at which evidence of successful reproduction was observed in the 1-2 years following translocation. -->

@@

@@

We then included demographic stochasticity (i.e., probabilistic births and deaths of individuals in a population) and environmental variability in the probability of juvenile recruitment to the adult class $\omega$ (the transition that we expect to be the most subject to environmental stochasticity in the presence of Bd) into our model described above and used it to predict long-term population viability.
Our model predicted that, following a single introduction of 40 adult individuals into a population and using the baseline value for $\omega$ of 0.3, all six populations with the highest adult survival probabilities had 50-year extinction probabilities less than 0.5 (0 - 0.32; \@fig-viability B).
This indicates strong potential for long-term persistence in the presence of Bd and environmental variability in recruitment (@fig-viability B).
When we decreased $\omega$ from 0.3 to our most conservative estimate of 0.15, 50 year extinction probability remained less than 0.5 for three populations and was 0.79 - 0.94 for three additional populations (@fig-viability B).
In contrast, for the six populations where $\sigma_{A_R} < 0.5$, extinction probability over 50 years was always predicted to be 1.
To test the validity of our model predictions, we demonstrated that our stochastic model could describe the general recovery trajectory of our translocated population with the longest survey history (@fig-viability-supp B; population 70550, surveyed for 16 years).

In summary, our model demonstrates that given observed yearly adult survival probabilities of translocated frogs, 50% of our translocated populations have a high probability of population growth and long-term viability in the presence of Bd.
This is likely a conservative estimate because there is evidence that naturally-recruited adults have higher survival probability than translocated adults (@fig-compare_surv_probs SI), but we considered these probabilities to be equal in all but three of our populations where we had sufficient data to distinguish these different probabilities.
Finally, five of the seven populations translocated after 2013 were predicted to have long-term viability, compared to only one of five populations translocated prior to 2013.
Thus, there is some trend in our improved ability to select appropriate habitat to promote long-term viability of translocated, resistant populations.

@@

@@

We provide three lines of evidence that support evolutionary rescue as a primary mechanism of recovery in MYL frog systems.
First, our genomic analyses comparing naive and recovering MYL frog populations found notable differences in allele frequencies between adjacent Bd-naive and recovering <!--# Check on using "persistent" vs "recovering" vs "recovered" --> populations, consistent with strong selection.
Importantly, some of the regions of the genome that showed signatures of selection are associated with known cellular and immunological mechanisms contributing to disease resistance in other species generally [@robledo2018; @robledo2020] and amphibians specifically [@rosenblum2008a; @ellison2014] (for more information, see SI X).
For example, we identified a \~940kb region on contig 8 that shows evidence of selection in recovering population and contains eight genes associated with either the MHC Class I or Class III regions.
These results corroborate many studies linking MHC genes to Bd survival [e.g., @savage2011; @bataille2015].
Broadly, our genomic analyses imply that selection and rapid evolutionary change has contributed to survival in MYL frogs in the presence of Bd.
<!--# For brevity in the discussion, I tried to a pull a few key bullets out of Allie's text. Allie and Bree: Edit and expand as you see fit. -->

The second line of evidence that evolutionary rescue is a strong candidate mechanism for recovery is that repeated translocations and long-term monitoring of naturally recovered populations have demonstrated that resistance phenotypes are maintained beyond the natal environment in the presence of Bd (as reflected by the high year-one adult survival in 6 out of 12 translocated populations).
Moreover, resistance has a quantifiable phenotype in this system -- resistant, persisting populations have Bd infection intensities below the level known to cause significant host mortality [@vredenburg2010, @knapp2016].
Translocated populations maintain infection intensities below this threshold of infection intensity such that infection intensity is not a significant predictor of adult survival following translocations to new environments.
Third and finally, high adult survival in the presence of Bd persists into subsequent generations in non-natal environments, suggesting that resistance is heritable to some degree.
Although previous studies have documented signatures of selection in hosts following disease-induced declines [@epstein2016, @gignoux-wolfsohn2021], our study is unique in that these three lines of evidence provide a strong case that evolved resistance begets population recovery in both natal and non-natal habitats, consistent with evolutionary rescue.

@@

Additionally, we identified a region on contig 19 that had much lower in the recovering group and consisted of one gene -- interferon-induced very large GTPase 1-like (GVINP1).
Interferons boost immune system processes and previous studies have linked interferon related genes to persisting post-Bd [@bryne2020] or increased expression when exposed to Bd in the lab [@rosenblum2008a; @ellison2014].
Variants in the specific gene identified here, GVINP1, explained \~20% of the resistance to Amoebic Gill Disease in Atlantic Salmon, a finding confirmed via expression analysis [@robledo2018; @robledo2020].
Furthermore, a key difference in susceptibility between populations of alpine tree frogs (*Litoria verreauxii alpina*) was the lower expression of IFN-induced very large GTPase in the more resistant, long-exposed population and upregulation of this same gene in more susceptible populations [@grogan2018].
Our study therefore contributes to the growing body of evidence linking this gene to disease resistance in general, and specifically to Bd survival in amphibians.
Finally, we found evidence of high FST and higher in recovering populations in the complement genes C6 and C7 <!--# It is clear why we talk about the other two regions (the ones highlighted in figure 2) and RIN3 - see comments above. But I guess the only reason this is interesting is because it is a complement region and it shows balancing rather than directional selection. If it feels like we can't justify discussing it we can cut it and just focus on the other two.  -->.
This could indicate that balancing selection is acting in this region of the genome, favoring a diverse set of alleles in the recovering group which has been shown for C6 in humans [@soejima2005].
Overall, key outlier genes in our dataset link directly to experimental and observational studies of immune responses to pathogens -- and often specifically to Bd responses in amphibians -- adding confidence to our conclusion that selection and rapid evolutionary change has contributed to survival in MYL frogs.

@@

The parameters in this model are yearly survival probability $\sigma_{\cdot}$ ("$\cdot$" indicates a particular state variable), probability that a female frog reproduces in a given year $p_F$, number of eggs produced by a female frog in a year that successfully hatch $F$, probability of a first-year tadpole remaining as a tadpole $p_{L1}$, probability of a second-year tadpole remaining as a tadpole $p_{L2}$, and probability of a first-year juvenile remaining as a juvenile \$p\_{J1}\$., and the probability of a juvenile successfully recruiting to an adult $\omega$.

@@

![A population viability analysis of translocated MYL frogs that quantifies the long-run growth rate and extinction probability of frog populations: (A) Predicted long-run growth rate $\lambda$ for different values of yearly adult survival probability $\sigma_{A_R}$ and probability of recruitment $\omega$, given the parameterized, deterministic model for translocated frogs. Colored points show the predicted $\lambda$ values for the twelve translocated populations when probability of recruitment into the adult class is $\omega = 0.3$ (given by the dashed line). The red line shows where $\lambda = 1$. (B) Predicted 50-year extinction probabilities of the 12 translocated populations given demographic stochasticity, environmental variability in $\omega$, and different mean values of $\omega$. The dashed line shows where $\omega = 0.3$, for comparison to (A). There are 6 lines at extinction probability = 1, 5 of which (70414--70619) are hidden beneath the line for 70505.](figures/pop_viability_figures_for_manuscript.jpg){#fig-viability}

![Sensitivity analysis and empirical validation of the stage-structured MYL frog model. (A) Elasticity of $\lambda$ with changes in three parameters: $\sigma_{A_R}$ (yearly survival probability of naturally recruited adults), $\omega$ (yearly probability of successful recruitment from juveniles to adults), and *F (*number of eggs produced by a female frog in a year that successfully hatch). Elasticity is calculated at the default parameter values for each population and $\omega$ = 0.3. (B) 100 simulated trajectories (black lines) that most closely matched the observed abundance trajectory of adult amphibians at site 70550 (light orange).](figures/pop_viability_figures_for_supp.jpg){#fig-viability-supp}

@@

![Comparison of the average yearly adult survival probabilities for adults translocated to each of three sites versus adults that were naturally recruited at each site (as a result of reproduction by translocated frogs). In contrast to @fig-translocation-survival, these are not survival probabilities from the first year following translocation, but instead represent averaged survival probabilities across multiple years and cohort. Points are median estimates and error bars give the 95% uncertainty intervals around the estimates, accounting for yearly variation in survival probabilities. The large error bars for 70550 reflect the large within and among individual variability in yearly survival probability at this site (as also reflected in one-year survival probabilities in [@fig-translocation-survival]). All estimates are derived from the mrmr package.](figures/compare_surv_probs.jpg){#fig-compare_surv_probs}

@@

```{r}
#| label: tbl-viability-earlylate
#| tbl-cap: "Proportion of populations translocated early versus late in the study period (< 2013 or >= 2013, respectively) for which the probability of extinction over 50 years was 1 or < 1 (i.e., viability = 0 or 1, respectively)."
#| echo: FALSE
#| message: FALSE

viability_table <- read_csv(here::here("data", "clean", "frog_translocation_final.csv")) %>% 
  distinct(site_id, year) %>% 
  group_by(site_id) %>% 
  filter(year == min(year)) %>% 
  mutate(site_id = as.character(site_id),
         viability = if_else(site_id %in% c(70414, 70370, 70134, 70641, 70619, 70505), 0, 1),
         period = if_else(year < 2013, "early", "late")) %>% 
  group_by(period) %>% 
  count(period, viability) %>% 
  mutate(n_period = sum(n), 
         prop = (n / n_period)) %>% 
  mutate(across(prop, round, 3)) %>% 
  arrange(viability, period)

  knitr::kable(viability_table, "pipe", align = "lcccr", col.names = c("Period", "Viability category", "Number of translocations", "Total translocations", "Proportion of translocations")) 
```
